---
title: "scRNA_vs_sn_RNAseq_DE"
author: "Ricardo Silva"
date: "5/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval=FALSE, 
                      message=FALSE, 
                      warning=FALSE)
```

```{r CLEAR EVERYTHING, eval=FALSE, include=FALSE}
# unload all non-base packages
lapply(names(sessionInfo()$otherPkgs), function(pkgs)
  detach(
    paste0('package:', pkgs),
    character.only = T,
    unload = T,
    force = T
))

rm(list=ls())
```

```{r}
#install.packages('Seurat')
#devtools::install_github("immunogenomics/harmony")
```

```{r load libraries, message=FALSE, warning=FALSE, include=FALSE}
suppressPackageStartupMessages({
  library(patchwork)
  library(tidyseurat) # to be compatible with Seurat and tidyverse APIs
  library(tidyverse)
})

```

```{r set wd and theme}
folder_path <- getwd()
source(paste0(folder_path,"/Script/theme_publication.R"))
theme_set(theme_Publication_3())

```

```{r}
#increasing R memory
options(future.globals.maxSize = 20000 * 1024^2)
```

Useful links:
<https://nbisweden.github.io/workshop-scRNAseq/exercises>
<https://hbctraining.github.io/scRNA-seq_online/schedule/>

###################################
Exploring known cell type markers
####################################
<https://hbctraining.github.io/scRNA-seq_online/lessons/08_SC_clustering_quality_control.html>

Let’s go through and determine the identities of the clusters. To access the normalized expression levels of all genes, we can use the normalized count data stored in the RNA assay slot.

NOTE: Assay is a slot defined in the Seurat object, it has multiple slots within it. In a given assay, the counts slot stores non-normalized raw counts, and the data slot stores normalized expression data. Therefore, when we run the `NormalizeData()` function in the above code, the normalized data will be stored in the data slot of the RNA assay while the counts slot will remain unaltered. `min.cutoff` of *q9* means 9% of cells with the lowest expression of the gene will not exhibit any purple shading (completely gray).

```{r get data from DE analysis}
markers_all <- read_csv(paste0(folder_path,"/Data/markers_obj.integ_cs_vs_nc_res0.7.csv"))
markers_cells <- read_csv(paste0(folder_path,"/Data/markers_cells_obj.integ_cs_vs_nc_res0.7.csv"))
markers_nucleous <- read_csv(paste0(folder_path,"/Data/markers_nucleous_obj.integ_cs_vs_nc_res0.7.csv"))
```

```{r venndiagramm for each dataset}
# VennDiagram
# Load library
library(VennDiagram)

# Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Generate 3 sets of each analysis
set1 <- markers_all %>% 
            group_by(cluster) %>% 
            slice_max(avg_log2FC, n = 100) %>% 
            pull(gene)
set2 <-  markers_cells %>% 
            group_by(cluster) %>% 
            slice_max(avg_log2FC, n = 100) %>% 
            pull(gene)
set3 <-  markers_nucleous %>% 
            group_by(cluster) %>% 
            slice_max(avg_log2FC, n = 100) %>% 
            pull(gene)

# create a list
 x = list(genes_all = set1,
          genes_cells = set2 , 
          genes_nucleous = set3)

# Helper function to display Venn diagram
display_venn <- function(x, ...){
  grid.newpage()
  venn_object <- VennDiagram::venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}

# Chart
display_venn(x,
             category.names = c("genes all" , "genes cells" , "genes nucleous"),
             
             # Circles
              lwd = 2,
              lty = 'blank',
              fill = myCol,
             
              # Numbers
              cex = 1.2,
              fontface = "bold",
              fontfamily = "sans",
              
              # Set names
              cat.cex = 1,
              cat.fontface = "bold",
              cat.default.pos = "outer",
              cat.pos = c(-27, 27, 135),
              cat.dist = c(0.055, 0.055, 0.085),
              cat.fontfamily = "sans",
              rotation = 1)

# Using ggVennDiagram
library(ggVennDiagram)
ggVennDiagram(x, 
              label_alpha = 0,
              category.names = c("genes sn and sn" , "genes sc" , "genes sn")) +
  scale_fill_gradient(low="white",high = "lightcoral") +
  scale_color_manual(values = c("grey", "grey", "grey")) +
  ggtitle("Venn Diagram - top 100 Avg Log2FC")

# get the intersections (only from github)
process_region_data(Venn(x)) %>% 
   unnest(cols = item) %>%
   select(-component, -id, -count) %>% 
   group_split(name) %>%
   walk(~write_csv(x = .x, file = paste0(folder_path,"/Output/Data/DE/all/top_100_",unique(.x$name), ".csv")))

```

```{r venndiagramm for each cluster}
# make a list with the top 100 DE genes
markers_list <-  list(marker_all = markers_all %>% 
                                      group_by(cluster) %>% 
                                      slice_max(avg_log2FC, n = 100) %>% 
                                      mutate(analysis = "marker_all"),
                       markers_cells = markers_cells %>% 
                                      group_by(cluster) %>% 
                                      slice_max(avg_log2FC, n = 100) %>%
                                      mutate(analysis = "markers_cells"),
                       markers_nucleous = markers_nucleous %>% 
                                      group_by(cluster) %>% 
                                      slice_max(avg_log2FC, n = 100) %>%
                                      mutate(analysis = "markers_nucleous"))

# join in one big tabel
markers_all_analysis <- map2_df(markers_list, names(markers_list), ~ mutate(.x, analysis = .y)) 

# nest by cluster
cluster_nested <- markers_all_analysis %>% 
  select(analysis, cluster, gene) %>%
  group_by(analysis, cluster) %>% 
  nest() 

# get the clusters to set names and and use in other functions below
n_clusters <- cluster_nested %>% pull(cluster) %>% as.factor() %>% levels()

# map each cluster
markers_per_cluster <- map(n_clusters , ~cluster_nested %>% filter(cluster == .x)) 

# make set list to be used in the function
sets_per_cluster <- markers_per_cluster %>% 
  map( ~pull(.) %>% flatten(.)) %>% 
  # names of outer lists
  set_names(paste0("Cluster_",n_clusters)) %>%
  # Use one more map() because the names to set are the inner lists.
  map(~ set_names(.,c("marker_all", "markers_cells", "markers_nucleous")))

# map the function
plot.list.venn <- sets_per_cluster %>% map2(.,names(sets_per_cluster),~
                           ggVennDiagram(.x,
                                         label_alpha = 0,
                                         category.names = c("genes sc and sn" , "genes sc" , "genes sn")) +
                           scale_fill_gradient(low="white",high = "lightcoral") +
                           scale_color_manual(values = c("grey", "grey", "grey")) +
                           ggtitle(paste0("Venn Diagram - top 100 DE genes in ",.y))
                         )
# print each plot
walk2(paste0(folder_path,"/Output/Plots/venn_clusters/top_100_",names(sets_per_cluster),".tiff"), plot.list.venn, ggsave)

# get the intersections and write in csv

# make filepaths from list names, which are unique values of the y column
#file_out <- sets_per_cluster %>% map(.,~paste0(folder_path,"/Output/Data/DE/clusters/top_100_",names(sets_per_cluster),"_",names(.),".csv"))

# write csv table for each cluster and intersections
# map interactions for each cluster
tablelist.top100 <- sets_per_cluster %>% map(.,~
                           process_region_data(Venn(.)) %>%
                           unnest(cols = item)) %>%
  # join all tables into a big table with a new column based on each cluster 
  map2_df(., names(.), ~ mutate(.x, cluster = .y)) %>%
  # split by clulster and interaction
  group_by(cluster, name) %>% 
  group_split() %>%
  # walk("map") write_csv function through each table
  walk(. %>% write_csv(paste0(folder_path,"/Output/Data/DE/clusters/top_100_",unique(.$cluster),"_",unique(.$name),".csv")))

```

