---
title: "scRNA_vs_sn_RNAseq_DE"
author: "Ricardo Silva"
date: "5/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval=FALSE, 
                      message=FALSE, 
                      warning=FALSE)
```

```{r CLEAR EVERYTHING, eval=FALSE, include=FALSE}
# unload all non-base packages
lapply(names(sessionInfo()$otherPkgs), function(pkgs)
  detach(
    paste0('package:', pkgs),
    character.only = T,
    unload = T,
    force = T
))

rm(list=ls())
```

```{r}
#install.packages('Seurat')
#devtools::install_github("immunogenomics/harmony")
```

```{r load libraries, message=FALSE, warning=FALSE, include=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(clustree)
  library(data.table)
  library(RColorBrewer)
  library(sctransform)
  library(scran)
  library(org.Mm.eg.db)
  library(clusterProfiler)
  library(formattable)
  library(patchwork)
  library(tidyseurat) # to be compatible with Seurat and tidyverse APIs
  library(tidyverse)
})

```

```{r set wd and theme}
folder_path <- getwd()
source(paste0(folder_path,"/Script/theme_publication.R"))
theme_set(theme_Publication_3())

```

```{r}
#increasing R memory
options(future.globals.maxSize = 20000 * 1024^2)
```

Useful links:
<https://nbisweden.github.io/workshop-scRNAseq/exercises>
<https://hbctraining.github.io/scRNA-seq_online/schedule/>

###################################
Exploring known cell type markers
####################################
<https://hbctraining.github.io/scRNA-seq_online/lessons/08_SC_clustering_quality_control.html>

Let’s go through and determine the identities of the clusters. To access the normalized expression levels of all genes, we can use the normalized count data stored in the RNA assay slot.

NOTE: Assay is a slot defined in the Seurat object, it has multiple slots within it. In a given assay, the counts slot stores non-normalized raw counts, and the data slot stores normalized expression data. Therefore, when we run the `NormalizeData()` function in the above code, the normalized data will be stored in the data slot of the RNA assay while the counts slot will remain unaltered. `min.cutoff` of *q9* means 9% of cells with the lowest expression of the gene will not exhibit any purple shading (completely gray).

```{r get the objects}
#Merged.sc <- read_rds(paste0(folder_path,"/Data/Merged.sc.clean.cc.cyclone.clust.RRPS.rds"))
Obj.integrated <- read_rds(paste0(folder_path,"/Data/Obj.integrated.cluster.RRPS.rds"))

# bellow the code to get these data 
#markers <- read_csv(paste0(folder_path,"/Data/markers_obj.integ_cs_vs_nc_res0.7.csv"))
annotations <- read_csv(paste0(folder_path,"/Data/annotation.csv"))

# Assign identity of clusters (identity as louvain with resolution x.x)
Idents(object = Obj.integrated) <- "integrated_snn_res.0.7"
#Idents(object = Obj.integrated) <- "integrated_snn_res.0.8"
```

```{r split the seurat obj}
Idents(Obj.integrated) <- "orig.ident"
single_cell <- subset(Obj.integrated, ident = "P0s_cell_RNAseq")
rm(Obj.integrated)
single_cell 
```


Start with marker genes for kidney cells
```{r known cell markers}
# change the default assay to RNA
DefaultAssay(single_cell) <- "RNA"

# Normalize RNA data for visualization purposes
Obj.integrated <- NormalizeData(single_cell, verbose = FALSE)

# to label the clusters
Idents(Obj.integrated) <- "integrated_snn_res.0.7"

# marker genes for kidney cells
kid.genes <- read_csv(paste0(folder_path,"/Data/kidney_marker_genes.csv"))

# add new column to be use in the plot titles
kid.genes <- kid.genes %>% mutate(title = case_when(str_detect(cell_type, "stromal") ~ "Stromal Cells",
                                       str_detect(cell_type, "imune") ~ "Imune system cells",
                                       str_detect(cell_type, "PT") ~ "Proximal Tubules",
                                       str_detect(cell_type, "CD_UB") ~ "UB tips and Collecting duct",
                                       str_detect(cell_type, "Macrophage") ~ "Macrophage",
                                       str_detect(cell_type, "CM") ~ "Cap Mesenchyme",
                                       str_detect(cell_type, "Epithelial_cells") ~ "Epithelial cells",
                                       str_detect(cell_type, "endothelial_cells") ~ "Endothelial cells",
                                       str_detect(cell_type, "Renal_Vesicle_IM") ~ "Renal Vesicle",
                                       str_detect(cell_type, "PD") ~ "Podocytes",
                                       str_detect(cell_type, "mesangial_cells") ~ "Mesangial cells",
                                       str_detect(cell_type, "renin-producing_cells") ~ "Renin-producing cells",
                                       str_detect(cell_type, "Distal_Metanephron") ~ "Distal Metanephron",
                                       str_detect(cell_type, "LOH") ~ "Loop of Henle",
                                       str_detect(cell_type, "DT") ~ "Distal Tubule")
                     )

# cheking
kid.genes %>% filter(str_detect(cell_type, "UB")) %>% pull(gene)

```

```{r table and resolution}
# cell types and clusters
cluster_names <- tibble(Cell_Type = c("CM", "PD", "PT", "LOH", "Stromal cells", "Endothelial cells", "Immune cells", "CD and UB"),
       Clusters_res.0.7 = c("2, 6", "3", "10, 17, 1", "8", "11, 4, 15", "0, 14", "18", "5, 9, 12"),
    #   Clusters_res.0.9 = c("1,9", "13, 14", "0, 5, 21", "6", "7, 11, 16, 17", "18, 4", "22", "3, 8, 15, 19")
       ) %>%
  separate_rows(., Clusters_res.0.7) %>% 
  mutate(cluster = as.factor(Clusters_res.0.7))

```


```{r}
# number of clusters
Idents(Obj.integrated) %>% n_distinct() # 21 clusters

# Rename all identities
seurat_integrated <- RenameIdents(object = Obj.integrated, 
                               "0" = "Endothelial cells",
                               "1" = "Proximal Tubules",
                               "2" = "Cap Mesenchyme",
                               "3" = "Podocytes",
                               "4" = "Stromal cells",
                               "5" = "UB and Collecting Ducts",
                               "6" = "Cap Mesenchyme",
                               "8" = "Loop of Henle",
                               "9" = "UB and Collecting Ducts",
                               "10" = "Proximal Tubules",
                               "11" = "Stromal cells",
                               "12" = "UB and Collecting Ducts",
                               "13" = "Podocytes",
                               "14" = "Endothelial cells",
                               "15" = "Stromal cells",
			            			       "17" = "Proximal Tubules", 
                    			     "18" = "Immune cells")
```


######################
Marker identification
######################
<https://hbctraining.github.io/scRNA-seq_online/lessons/09_merged_SC_marker_identification.html>
<https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_05_dge.html>

First get annotations 
<https://hbctraining.github.io/scRNA-seq_online/lessons/fetching_annotations.html>
```{r annotation}
# Connect to AnnotationHub
library(AnnotationHub)
ah <- AnnotationHub()

# Access the Ensembl database for organism
ahDb <- query(ah, 
              pattern = c("Mus musculus", "EnsDb"), 
              ignore.case = TRUE)

# Acquire the latest annotation files
id <- ahDb %>%
        mcols() %>%
        rownames() %>%
        tail(n = 1)

# Download the appropriate Ensembldb database
edb <- ah[[id]]

# Extract gene-level information from database
annotations <- genes(edb, 
                     return.type = "data.frame")

# Select annotations of interest
annotations <- annotations %>%
        dplyr::select(gene_id, gene_name, seq_name, gene_biotype, description)

# save
# write_csv(annotations, paste0(folder_path,"/Data/annotation.csv"))

```

Identification of all markers for each cluster:

Now we wil compare each cluster against all others and outputs the genes that are DE/present
Good to improve confidence in hypothesized cell types
NOTE: `FindAllMarkers` function is recommended for when evaluating a single sample group/condition

```{r all markers}
# Find markers for every cluster compared to all remaining cells, report only the positive ones
markers_cells <- FindAllMarkers(object = Obj.integrated, 
                          test.use = "wilcox",
                          min.pct = 0.25,
                          min.diff.pct = 0.25,
		                      max.cells.per.ident = 50,
                          only.pos = TRUE, # keep only the positive changes
                          logfc.threshold = 0.25,
		                      assay = "RNA") #minimum log fold change (natural log) for average expression of gene

# join annotations
markers_cells <-markers_cells %>%
    left_join(y = annotations %>% distinct(gene_name, description),
               by = c("gene" = "gene_name")) %>% 
    left_join(cluster_names)

# number of genes per cluster
markers_cells %>% group_by(cluster) %>% summarise(n_gene = n())

# save
#write_csv(markers_cells, paste0(folder_path,"/Data/markers_cells_obj.integ_cs_vs_nc_res0.7.csv"))
#markers_cells <- read_csv(paste0(folder_path,"/Data/markers_cells_obj.integ_cs_vs_nc_res0.7.csv"))

# top 10 and 5 markers per cluster
top10_cells <- markers_cells %>% group_by(cluster) %>% 
  slice_max(avg_log2FC, n = 10) 
top5_cells <- markers_cells %>% group_by(cluster) %>% 
  slice_max(avg_log2FC, n = 5) 

```

```{r bar plot top_genes}
# number of cluster
clusters.top <- top10_cells %>% distinct(cluster) %>% pull()
name.cell <-  top10_cells %>% distinct(Cell_Type) %>% pull()

# plot for each cluster the top 10 marker genes based on the avg log2 FC
plot.list.marker.genes <- clusters.top %>% map2(.,name.cell, ~
      top10_cells %>% filter(cluster == .x) %>%
          mutate(gene = fct_reorder(gene,avg_log2FC)) %>%
      ggplot(aes(x = gene, y = avg_log2FC)) +
          geom_bar(stat = "identity", fill="#f68060", alpha=.6, width=.4) +
          geom_text(aes(label = description), y=0.1, hjust=0) +
          coord_flip() +
          ggtitle(glue::glue("{.y}: Cluster {.x} vs. rest - P0 single nucleous")) +
          theme(plot.title = element_text(hjust = 0.5)) +
          labs(x = "Marker genes (annotations)",
               y = "avg log2 Fold-Change")

)

plot.list.marker.genes[[2]]

map2(paste0(folder_path,"/Output/Plots/only_cells/top_gene_markers/cluster_",clusters.top,"_vs_rest.tiff"), plot.list.marker.genes, ggsave)

```

```{r heatmap and dotplot}
# Set the identity as louvain with resolution 0.5
sel.clust = "integrated_snn_res.0.7"

# create a scale.data slot for the selected genes
Obj.integrated <- ScaleData(Obj.integrated, 
                            features = as.character(unique(top5cells$gene)), 
                            assay = "RNA")
DoHeatmap(Obj.integrated, 
          features = as.character(unique(top5cells$gene)), 
          group.by = sel.clust,
          assay = "RNA")

DotPlot(Obj.integrated, 
        features = rev(as.character(unique(top5cells$gene))), 
        group.by = sel.clust,
        assay = "RNA") + coord_flip()
```

```{r plot staked vnplot}
# select all cells in cluster 1
features1 <- top5_cells %>% distinct(gene) %>% pull() %>% head(10)

# to stack vnplots -> stack=T , flip = T
VlnPlot(Obj.integrated, 
        features = features1, 
        stack=T , 
        flip = T) + 
  NoLegend() + ggtitle('Single Cells')

VlnPlot(Obj.integrated, features = features1, stack=T , flip = T) & 
  theme(legend.position = "bottom",
        legend.direction = "horizontal")
```



Other kinds of Find Markers

Identification of conserved markers for each cluster

It looks for genes that are diff expressed/present within each condition first, and then reports those genes that are conserved in the cluster across all conditions
Useful with more than one condition to identify cell type markers that are conserved across conditions.
`FindConservedMarkers` function separates genes by sample group/condition. Then, it performs DE testing for a single specific cluster against all the others
The output is matrix containing a ranked list of putative markers listed by gene ID for the cluster we specified, and associated stats. Below some columns showed in the matrix:
- condition_avg_logFC: average log fold change for condition. Positive values indicate that the gene is more highly expressed in the cluster.
- condition_pct.1: percentage of cells where the gene is detected in the cluster for condition
- condition_pct.2: percentage of cells where the gene is detected on average in the other clusters for condition
- condition_p_val_adj: adjusted p-value for condition, based on bonferroni correction using all genes in the dataset, used to determine significance

NOTE: When looking at the output, we suggest *looking for markers with large differences in expression between `pct.1` and `pct.2` and larger fold changes*. For instance if `pct.1` = 0.90 and `pct.2` = 0.80, it may not be as exciting of a marker. However, if `pct.2` = 0.1 instead, the bigger difference would be more convincing. Also, of interest is if the majority of cells expressing the marker is in my cluster of interest. If `pct.1` is low, such as 0.3, it may not be as interesting. Both of these are also possible parameters to include when running the function, as described above.

NOTE: !!!!! The function fail because some clusters that do not have enough cells for a particular group !!!!
run `FindAllMarkers()` instead (Above).
```{r conserved marker}
## DO NOT RUN ##

# set our default assay. We want to use the normalized data, not the integrated data
DefaultAssay(Obj.integrated) <- "RNA"

# gene annotation information
annotations <- read_csv(paste0(folder_path,"/Data/annotation.csv"))

# Create function to get conserved markers for any given cluster
get_conserved <- function(cluster, group){
  FindConservedMarkers(Obj.integrated,
                     ident.1 = cluster, # specify a single cluster 
                     grouping.var = "orig.ident", 
                     only.pos = TRUE,
		                 min.diff.pct = 0.25,
		                 max.cells.per.ident = 50,
                     min.pct = 0.25, # detect at a minimum percentage in either of the two groups of celss
		                 logfc.threshold = 0.25) %>%
    rownames_to_column(var = "gene") %>%
    left_join(y = annotations %>% distinct(gene_name, description),
               by = c("gene" = "gene_name")) %>%
    cbind(cluster_id = cluster, .)
}

# Iterate function across desired clusters
conserved_markers <- map_dfr(c(levels(Obj.integrated)), get_conserved)

```


```{r another annotation}
# Load additional packages
library(enrichR)

# Check available databases to perform enrichment (then choose one)
enrichR::listEnrichrDbs() %>% filter(if_any(everything(), ~ str_detect(. ,"Mouse"))) 

# Perform enrichment
enrich_results <- enrichr(genes = DGE_cell_selection$gene[DGE_cell_selection$cluster == 
    "Covid"], databases = "KEGG_2019_Mouse")[[1]]
```




